---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
options(digits = 4, width = 120)
```

# rpxl: Read password-protected Excel files
<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![R-CMD-check](https://github.com/patrickbarks/xltest17/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/patrickbarks/xltest17/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

Uses the Python library
[msoffcrypto-tool](https://pypi.org/project/msoffcrypto-tool/) to decrypt
password protected Excel files (either .xlsx or .xlsb), which can then be read
into R with e.g. the [readxl](https://readxl.tidyverse.org/) or
[readxlsb](https://github.com/velofrog/readxlsb) R packages.

## Installation

Install from GitHub with:

```{r, eval=FALSE}
# install.packages("remotes")
remotes::install_github("epicentre-msf/rpxl")
```

Next, make sure Python is installed on your system. You can use the
[reticulate](https://rstudio.github.io/reticulate/index.html) R package for
this, which will already be installed alongside rpxl.

```{r, eval=FALSE}
reticulate::py_available(initialize = TRUE)
```

If you don't already have Python installed, it can be downloaded from
www.python.org/downloads, or installed directly from R using the
`install_python()` function in reticulate.

```{r, eval=FALSE}
reticulate::install_python()
```

Finally, use the `install_rpxl()` function to install the required Python
dependencies.

```{r, eval=FALSE}
library(rpxl)
install_rpxl()
```



## Usage

```{r}
library(rpxl)
```

Read password-protected _.xlsx_ files with function `rpxlsx()`, a wrapper to
`read_xlsx()` in the [readxl](https://readxl.tidyverse.org/) package.

```{r}
path_xlsx <- system.file("extdata", "xltest.xlsx", package = "rpxl")
rpxlsx(path_xlsx, password = "1234")
```

Read password-protected _.xlsb_ files with function `rpxlsb()`, a wrapper to
`read_xlsb()` in the [readxlsb](https://github.com/velofrog/readxlsb) package.
Note that `read_xlsb()` requires the worksheet to be explicitly specified
(either with argument 'sheet' or 'range'), whereas `read_xlsx()` defaults to
sheet 1.

```{r}
path_xlsb <- system.file("extdata", "xltest.xlsb", package = "rpxl")
rpxlsb(path_xlsb, sheet = 1L, password = "1234")
```

To decrypt a password-protected Excel file and save it under a different
filename (without immediately reading it into R), you can use the function
`rpxl_decrypt()`. The function returns the path to the decrypted file, which
defaults to a temporary file created with `tempfile()`, but could alternatively
be user-specified.

```{r}
path_decrypted <- rpxl_decrypt(path_xlsx, password = "1234")
```

The decrypted file can then be read into R in a separate step.

```{r}
readxl::read_xlsx(path_decrypted)
```

You may want to explicitly delete the decrypted file afterwards, if its contents
are sensitive. Otherwise, if it's a temporary file created with `tempfile()`, it
will be automatically deleted when the current R session is closed.

```{r}
file.remove(path_decrypted)
```

